---
title: Stan code for the seedling survival model
author: Masatoshi Katabuchi
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
#rmdformats::material
  html_document:
    fig_caption: yes
    theme: spacelab #readable #sandstone #spacelab #flatly
    highlight: pygments #tango #kate
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
    toc_float:
      collapsed: TRUE
      smooth_scroll: TRUE
      toc: true
---

```{r, echo = F}
rm(list = ls())
```

In this case study I will demonstrate how seedling survival models are implemented with Stan using a simple example. Basically you can apply the same code to your seedling data. This example has reduced numbers of samples and parameters than what we are trying to do but it still takes a relatively long time to run.

# Model overview

Survival (*s*) of an individual seedling *i*, of species *j*, in plot *l*, in census *t* is model as:

$$
S_{ijlt} \sim Bernoulli(p_{jlt})
$$

$$
logit(p_{jlt}) = \boldsymbol{\beta_{j}} \cdot \boldsymbol{X} + \phi_l + \tau_t
$$

where $\boldsymbol{\beta_{j}} = \left[\beta_{0j}, \beta_{1j}, \ldots, \beta_{kj} \right]$ is the vectors of coefficients at the species-level; *k* is the number of predictors for an individual seedling; $\boldsymbol{X} = \left[X_{0}, X_{1}, \ldots, X_{k} \right]$ is the predictors for an individual seedling; $\phi_l$ is a random effect for seedling plot, where $\phi_l \sim \mathcal{N}(0,\sigma_{\phi})$; and $\tau_t$ is a random effect for census, where $\phi_l \sim \mathcal{N}(0,\sigma_{\phi})$ (Note that $\cdot$ denotes dot product).

In the species-level regression, the vector of coefficients ($\beta_{0-k}$) of each species j can be assumed have a multivariate normal distribution (MVN);

$$
\boldsymbol{\beta_j} \sim  \mathcal{MVN}(\gamma_{n0} + \gamma_{n1} \cdot WP_j, \bf {\Sigma_{\beta}})
$$

where $\gamma_{n0-1}$ are the vectors of coefficients; $WP_j$ is the water potential for species *j*; and $\bf \Sigma_{\beta}$ is the covariance matrix.

The covariance matrix $\bf \Sigma_{\beta}$ can be decomposed as ${\bf \Sigma} = {\rm diag}({\bf \sigma}){\bf \Omega}{\rm diag}({\bf \sigma}) = {\rm diag}({\bf \sigma}){\bf LL}\prime {\rm diag}({\bf \sigma})$ using a Cholesky decomposition (See [Alvarez et al. 2014](https://arxiv.org/abs/1408.4050), [Katabuchi et al. 2017](https://www.biorxiv.org/content/10.1101/116855v1.abstract) and [this post](https://stla.github.io/stlapblog/posts/stanlkjprior.html)). See the model code below for the priors.


```{r}
writeLines(readLines("model.stan"))
```


# Dummy data

We generate dummy data with 20 species, 100 seedlings per species, 10 plots and 5 censuses. In total we have 2000 seedlings.

```{r}
n_sp <- 20
sp_name <- paste0("sp", LETTERS[1:n_sp])
n_para <- 5
n_tree <- 100
n_plot <- 10
n_census <- 5
n_sample <- n_sp * n_tree
```


```{r}
set.seed(12345)
library(tidyverse)

WP <- rnorm(n_sp, 0, 1) %>% round(3)
WP_dat <- tibble(sp = 1:n_sp, WP)
U <- cbind(Intercept = 1, WP)
rownames(U) <- sp_name
U
```
U is the species level predictor.

```{r}
gamma0 <- rnorm(n_para, 0, 1) %>% round(3)
gamma1 <- c(0, 0.8, -0.4, 0, 0)
gamma <- rbind(gamma0, gamma1)

colnames(gamma) <-  c("Intercept",
                 "CONS",
                 "HETS",
                 "CONBA",
                 "HETBA")

u_gamma <- U %*% gamma
rownames(u_gamma) <- sp_name
colnames(u_gamma) <- c("Intercept",
                 "CONS",
                 "HETS",
                 "CONBA",
                 "HETBA")

#u_gamma %>% round(3)

## not using MVN for simplicity
R <- matrix(rnorm(n_para * n_sp, 0, 0.2), ncol = n_para)
beta <- u_gamma + R

sigma_phi <- 0.1
sigma_tau <- 0.2

phi <- rnorm(n_plot, 0, sigma_phi)
tau <- rnorm(n_census, 0, sigma_tau)

```

Some paratemers include:


- $\bf \gamma$

```{r, echo = F}
gamma %>% t
```


- $\sigma_\phi$ = 0.1
- $\sigma_\tau$ = 0.2
- $\sigma$ in $\bf \Sigma$ = 0.2 and $\rho$ = 0
- $\bf \beta$ =

```{r, echo = F}
beta %>% round(3)
```

In this example, we only have 5 variables for the individual predictor, but we could add more effects (e.g., soil, season and rainfall).

```{r}

#plot(u_gamma, beta)
suv_mat <- matrix(numeric(n_sp * n_tree), nrow = n_sp)
X <- NULL
for (sp in 1:n_sp) {
  X0 <- matrix(rnorm(n_para * n_tree, 0, 1), nrow = n_para)
  X0[1,] <- 1
  suv_mat[sp, ] <- beta[sp, ] %*% X0
  X <- cbind(X, X0)
}

rownames(suv_mat) <- sp_name

p0 <- suv_mat %>%
  t %>%
  as.numeric

plot <- sample(1:n_plot, n_sample, replace = T)
census <- sample(1:n_census, n_sample, replace = T)

p00 <- p0 + phi[plot] + tau[census]

logistic <- function(z) {
  1 / (1 + exp(-z))
}

p <- logistic(p00)
suv <- rbinom(n_sample, 1, p)

dat <- tibble(suv, p = p %>% round(3), p00 = p00 %>% round(3)) %>%
 mutate(plot = plot) %>%
 mutate(census = census) %>%
 mutate(sp = rep(1:n_sp, each = n_tree)) %>%
 full_join(., WP_dat, by = "sp")

DT::datatable(dat)

```

A generated survival data set looks like this.

```{r}

hist_dat <- tibble(S = suv,
                   p = p,
                   `logistic(p)` = p00) %>%
  gather(dat, val, 1:3)

ggplot(hist_dat, aes(x = val)) +
  geom_histogram() +
  facet_wrap(~ dat, scale = "free")

```

# Model

We then use the Stan to fit the data.

```{r, cache=TRUE, comment=NA}
library(rstan)
rstan::rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores()) # Run on multiple cores

list_dat <- list(N = nrow(dat),
                 J = n_sp,
                 K = n_para,
                 S = n_plot,
                 T = n_census,
                 L = 2,
                 suv = suv,
                 plot = plot,
                 census = census,
                 sp = dat$sp,
                 x = t(X),
                 u = U)

fit <- stan(file = "model.stan",
            data = list_dat,
            iter = 2000,
            warmup = 1000,
            thin = 1,
            chains =  4,
            refresh = 500,
            control = list(adapt_delta = 0.9, max_treedepth = 20))

```

The resulting estimates well recovered the original values (e.g., $\bf \gamma$, $\sigma_{\tau}$ and $\sigma_{\phi}$) `r emo::ji("beer")`



```{r}
print(fit, pars = c("gamma", "sigma_phi", "sigma_tau"))
```

```{r}
gamma %>% t
```

- $\sigma_\phi$ = 0.1
- $\sigma_\tau$ = 0.2

# Computing Environment

```{r}
writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))
devtools::session_info("rstan")
```
