---
fontsize: 12pt
geometry: margin=1in
link-citations: yes
csl: templates/ecology-letters.csl
bibliography: templates/seedling.bib
crossref:
  fig-title: Fig.
  fig-prefix: Fig.
  eq-prefix: Eq.
format:
  docx:
    toc: false
    number-sections: false
    highlight-style: github
    html-math-method: katex
  html:
    theme: spacelab #readable #sandstone #spacelab #flatly
    toc: true
    toc-depth: 2
    toc-title: Contents
    self-contained: true
    smooth-scroll: true
    highlight-style: github
  pdf:
    toc: false
    keep-tex: true
    pdf-engine: pdflatex
    include-in-header:
      text: |
        \usepackage{xr}
        \usepackage[default]{sourcesanspro}
        \usepackage{sourcecodepro}
        \usepackage{lineno}
        \linenumbers
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(patchwork)
library(here)
library(kableExtra)
```

```{r}
source(here("R", "stan.R"))
source(here("R", "render.R"))
```

```{r, include=FALSE}
s_table("dry_traits_diagnostics")
s_table("wet_traits_diagnostics")
s_table("dry_abund_diagnostics")
s_table("wet_abund_diagnostics")
s_fig("site_info")
s_fig("rainfall")
s_fig("cc_line")
```

# Model

We modeled the seedling survival for the dry and rainy seasons separately.
Since the effect of tree neighbors on seedling survival is nonlinear on a logistic scale [@Detto2019], we performed a grid-search for the scaling parameter *c* between 0 and 1 in 0.01 increments that maximized the likelihood of the following survival model,

$$
\mathrm{logit}(p_i) = b_0 + b_1 Z_{1i}^c + b_2 Z_{2i}^c,
$$

where $p_i$ is the individual survival probability in the *i*th census interval, and $Z_1$ and $Z_2$ are distance-weighted sums of basal areas of conspecifics and heterospecifics respectively.
We found that $c$ = 0.27 for the dry season and $c$ = 0.24 for the rainy season were the best estimates for our dataset (`r s_fig("cc_line")`).

We built Bayesian hierarchical models that include variation among species in the effects of conspecific and heterospecific neighbours, and rainfall on survival.
Survival ($s$) of seedling record *i* of individual *m* for species *j* in census *t* in plot *p* was modeled using the Bernoulli distribution ($\mathcal{B}$):

$$
s_{i,j,m,t,p} \sim \mathcal{B}(p_{i, j, t, p}),
$$

$$
\mathrm{logit}(p_{i,j,t,p}) = \boldsymbol{x_{i}} \cdot \boldsymbol{\beta_{j}} + \phi_p + \omega_t,
$$

where $\boldsymbol{\beta_{j}} = \left[\beta_{1,j}, \beta_{2,j}, \ldots, \beta_{K,j} \right]$ is the coefficient row *K*-vector for species *j*,
*K* is the number of predictors for an individual seedling,
$\boldsymbol{x_i} = \left[x_{i,1},x _{i,2}, \ldots,x_{i,K} \right]$ is the vector of predictors of size *K* for an individual seedling,
$\phi_p$ is the random effect for seedling plots, and
$\omega_t$ is the random effect for different censuses
(note that $\cdot$ denotes dot product).
The set of predictor variables ($\boldsymbol{x_i}$) includes
intercepts,
log of seedling heights,
rainfall,
densities of conspecific (*ConS*) and heterospecific (*HetS*) seedlings,
densities of conspecific (*ConA*) and heterospecific (*HetA*) trees that are scaled by 0.27 for the dry season or 0.24 for the rainy season,
and the interactions of rains with *ConS*, with *HetS*, with *ConA* and with *HetA*.
We also considered models with phylogenetic density effects instead of heterospecific density effects.

In the species-level regression, the row vector of coefficients ($\boldsymbol{\beta_{1-K}}$) of each species *j* were assumed to have a multivariate normal distribution through the Cholesky factorization (@StanDevelopmentTeam2021; `r s_append("method")`),

$$
\boldsymbol{\beta_j} = \boldsymbol{\gamma_k} \cdot \boldsymbol{u_j} + \boldsymbol{\Sigma}
$$

where $\boldsymbol{u_{j}} = \left[u_{1,j}, u_{2,j}, \ldots, u_{L,j} \right]$ is the row vector of predictors of size *L* for species *j*,
*L* is the number of predictors for each species (i.e., the number of traits including an intercept),
$\boldsymbol{\gamma_k} = \left[\gamma_{k,1}, \gamma_{k,2}, \ldots, \gamma_{k,L} \right]$ is the coefficient *L*-vector for *k*th predictor in the individual-level regression, and
$\boldsymbol{\Sigma}$ is the covariance matrix of the multivariate normal distribution.
We modeled five different sets of species-level predictors separately.
The set of species-level predictor variables ($\boldsymbol{u_j}$) includes
1)
LDMC,
SDMC,
LA,
SLA,
Chl,
LT,
$\delta$C~13~,
C,
N, and
$\pi$~tlp~,
2) abundance,
3) basal area, and
4) abundance and basal area.
The row vector $\gamma_{k,1}$ represents the average effects of each individual-level predictor (e.g., *ConS*) across species,
whereas $\gamma_{k, l}$ ($l \ne 1$), represents the effects of the *l*-th individual-level predictor (e.g., SLA) on the variation in the strength of each individual-level predictor (e.g., variation in the strength of *ConS* among species).
To allow comparisons among parameter estimates, the individual-level predictors ($\boldsymbol{x_i}$) and the species-level predictors ($\boldsymbol{u_j}$) were scaled to a mean of 0 and standard deviation of 1 within each season and across species, respectively.

Posterior distributions of all parameters were estimated using the Hamiltonian Monte Carlo algorithm (HMC) implemented in Stan [@Carpenter2017] using the weakly-informative priors [@Gelman2008].
Convergence of the posterior distribution was assessed with the Gelman-Rubin statistic with a convergence threshold of 1.1 [@Gelman2013] and effective sample sizes > 400 [@Vehtari2021] for all parameters.
All statistical analyses were conducted in R version 4.2.1 [@RCoreTeam2022] using the R package *targets* version 0.13.5 for workflow management [@Landau2021].
For further details, see `r s_append("method")`.
Codes and datasets are available at
https://github.com/mattocci27/trait-ndd
and
XXXX,
respectively.

# References

<div id="refs"></div>

\newpage

```{r, include=FALSE}
dry_traits_gamma <- read_csv(here("data", "dry_het_intrain2_traits_gamma.csv"))
dry_traits_gamma |>
  filter(q2.5 * q97.5 > 0)
dry_traits_gamma |>
  filter(trait_name == "intercept")
```

```{r, include=FALSE}
gen_si_tab2 <- function(data) {
  data2 <- data |>
    mutate(pred_name = case_when(
      pred_name == "(Intercept)" ~ "Intercept",
      pred_name == "logh_s" ~ "ln Height",
      pred_name == "scon_s" ~ "ConS",
      pred_name == "acon_s_c" ~ "ConT",
      pred_name == "shet_s" ~ "HetS",
      pred_name == "ahet_s_c" ~ "HetT",
      pred_name == "sphy_s" ~ "PhyS",
      pred_name == "aphy_s_c" ~ "PhyT",
      pred_name == "rain_s" ~ "Rainfall",
      pred_name == "scon_s:rain_s" ~ "ConS $\\times$ Rainfall",
      pred_name == "acon_s_c:rain_s" ~ "ConT $\\times$ Rainfall",
      pred_name == "shet_s:rain_s" ~ "HetS $\\times$ Rainfall",
      pred_name == "ahet_s_c:rain_s" ~ "HetT $\\times$ Rainfall",
      pred_name == "sphy_s:rain_s" ~ "PhyS $\\times$ Rainfall",
      pred_name == "aphy_s_c:rain_s" ~ "PhyT $\\times$ Rainfall",
      pred_name == "scon_s:rain_s" ~ "ConS $\\times$ Rainfall",
      pred_name == "logh_s:rain_s" ~ "ln Height $\\times$ Rainfall",
    )) |>
    mutate(trait_name = case_when(
      trait_name == "intercept" ~ "Intercept",
      trait_name == "ldmc" ~ "LDMC",
      trait_name == "sdmc" ~ "SDMC",
      trait_name == "chl" ~ "Chl",
      trait_name == "c13" ~ "$\\delta \\mathrm{C_{13}}$",
      trait_name == "log_n" ~ "ln N",
      trait_name == "tlp" ~ "$\\pi_\\mathrm{{tlp}}$",
      trait_name == "log_la" ~ "ln LA",
      trait_name == "log_sla" ~ "ln SLA",
      trait_name == "log_lt" ~ "ln LT",
      trait_name == "log_ab" ~ "ln Abundance",
      )) |>
    mutate_if(is.numeric, round, 3) |>
    filter(q2.5 * q97.5 > 0) #|>
  data2 |>
    mutate(`90\\% CI` = paste0("[", q5, ", ", q95, "]")) |>
    mutate(`95\\% CI` = paste0("[", q2.5, ", ", q97.5, "]")) |>
    mutate(variable = paste0(
      "$\\gamma_{",
      str_split_fixed(data2$variable, ",|\\[|\\]" ,4)[,2],
      ",",
      str_split_fixed(data2$variable, ",|\\[|\\]" ,4)[,3],
      "}$")) |>
    rename(Parameter = variable) |>
    rename(Median = q50) |>
    rename(`Lower 2.5\\% CI` = q2.5) |>
    rename(`Lower 5\\% CI` = q5) |>
    rename(`Upper 95\\% CI` = q95) |>
    rename(`Upper 97.5\\% CI` = q97.5) |>
    rename(`Individual-level predictor` = pred_name) |>
    rename(`Species-level predictor` = trait_name) |>
    dplyr::select(Parameter, Median,
    `95\\% CI`,
    `Individual-level predictor`,
    `Species-level predictor`
    ) |>
    kbl(booktabs = TRUE, escape = FALSE, format = "pipe", longtable = TRUE) |>
    kable_styling(latex_options = c("striped", "scale_down", "HOLD_position", "repeat_header"))
}
```

# `r table("dry_traits_gamma")`

Posterior medians and 95% credible intervals (CIs) of the effects of species-level predictors (traits) on individual-level predictors ($\gamma$) in the dry seasons.
Only variables with credible intervals not overlapping 0 are shown.
The full results are available in `r s_table("dry_traits_gamma")`.
ConS/HetS: conspecific/heterospecific seedling density,
ConT/HetT: conspecific/heterospecific tree density (scaled by 0.27),
LDMC: leaf dry matter content,
SDMC: stem dry matter content,
$\delta$C~13~: stable carbon isotope composition,
LT: leaf thickness.

```{r}
gen_si_tab2(dry_traits_gamma)
```

\newpage

# `r table("wet_traits_gamma")`

Posterior medians and 95% credible intervals (CIs) of the effects of species-level predictors (traits) on individual-level predictors ($\gamma$) in the rainy seasons.
Conspecific (ConT) and phylogenetic (PhyT) tree densities are scaled by 0.24.
The full results are available in `r s_table("wet_traits_gamma")`.
$\delta$C~13~: stable carbon isotope composition,
Other details as for `r table("dry_traits_gamma")`.

```{r, include=FALSE}
wet_traits_gamma <- read_csv(here("data", "wet_phy_norain_traits_gamma.csv"))
wet_traits_gamma |>
  filter(q2.5 * q97.5 > 0)
wet_traits_gamma |>
  filter(trait_name == "intercept")
```

```{r}
gen_si_tab2(wet_traits_gamma)
```

\newpage

# `r fig("coef_plot")`

Standardized regression coefficients common to all species ($\gamma_{k,1}$) modeling the effects of
log of individual height (ln Height),
conspecific seedling density (ConS),
conspecific tree density (ConT),
heterospecific seedling density (HetS),
phylogenetic seedling density (PhyS),
heterospecific tree density (HetT),
phylogenetic tree density (PhyT),
rainfall, and
interactions between densities and rainfall for dry and rainy seasons.
Thick and thin lines indicate 90% and 95% credible intervals, respectively.
Circles show posterior means of coefficients.
Filled circles indicate significant effects and open circles indicate non-significance effects.
Positive $\gamma_{k,1}$ values indicate higher survival rates with increasing values of the predictors,
while negative $\gamma_{k,1}$ values indicate lower survival rates with increasing values of the predictors.
Note that predictors are scaled to a mean of 0 and standard deviation of 1 within each season.

![](`r here::here("figs/coef_trait.pdf")`)

\newpage

# `r fig("dry_contour_plot")`

Contour plots illustrating interactions between conspecific seedling densities and rainfall on seedling survival rates in the dry seasons and their association with functional traits:
(a) leaf dry matter content (LDMC),
(b) stem dry matter content (SDMC),
(c) leaf thickness (LT),
and (d) stable carbon isotope composition ($\delta$C~13~).
Colour scales show marginal estimated median predictions for each response variable across the full range of rainfall and upper 90% quantiles of conspecific seedling density (the distributions of conspecific seedling densities are positively skewed).
All other variables were set at their mean value for marginal predictions.
Low and high trait values indicate 25% and 75% quantiles of the trait distribution in the observed dataset, respectively.
Only variables with credible intervals not overlapping 0 are plotted.

![](`r here::here("figs/dry_het_intrain2_trait_suv_contour.pdf")`)

