---
fontsize: 12pt
geometry: margin=1in
link-citations: true
csl: templates/ecology-letters.csl
bibliography: templates/seedling.bib
crossref:
  fig-title: Fig.
  fig-prefix: Fig.
  eq-prefix: Eq.
format:
  docx:
    toc: false
    number-sections: false
    highlight-style: github
    html-math-method: katex
  pdf:
    toc: false
    keep-tex: true
    pdf-engine: pdflatex
    include-in-header:
      text: |
        \usepackage{xr}
        \usepackage[default]{sourcesanspro}
        \usepackage{sourcecodepro}
        \usepackage{lineno}
        \linenumbers
  html:
    theme: spacelab #readable #sandstone #spacelab #flatly
    toc: true
    toc-depth: 2
    toc-title: Contents
    self-contained: true
    smooth-scroll: true
    highlight-style: github
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(patchwork)
library(here)
library(kableExtra)
```

```{r}
source(here("R", "stan.R"))
source(here("R", "render.R"))
```

# Survival models

We first modeled seedling survival at the individual level to assess whether neighbor densities have different effects between the dry and rainy seasons and whether phylogenetic-weighted neighbor densities affect seedling survival differently than the heterospecific neighbor densities.
We found strong positive correlations between heterospecific neighbor densities and phylogenetic-weighted neighbor densities (Figure S5).
Thus, the effects of phylogenetic-weighted neighbor densities and heterospecific neighbors on seedling survival were comparable (Table S1).
We also detected seasonal interactions with neighbor densities (Table S1).
We will focus on heterospecific neighbor density in our subsequent trait-mediated survival models for each season separately based on these findings.
These preliminary models also incorporated a scaling parameter c to account for nonlinear effects on a logistic scale (see below).

# Traits and abundance mediate survival models

Since the effect of tree neighbors on seedling survival is nonlinear on a logistic scale [@Detto2019], we performed a grid search for the scaling parameter *c* between 0 and 1 in 0.01 increments that maximized the likelihood of the following survival model,

$$
\mathrm{logit}(p_i) = b_0 + b_1 Z_{1i}^c + b_2 Z_{2i}^c,
$$

where $p_i$ is the survival probability of the *i*th individual, and $Z_1$ and $Z_2$ are distance-weighted sums of basal areas of conspecifics and heterospecifics respectively.
We found the best estimates of *c* to be 0.27 for the dry season, 0.24 for the rainy season, and 0.26 for the combined analysis of both seasons (`r s_fig("cc_line")`).

We built Bayesian hierarchical models that include variation among species in the effects of conspecific and heterospecific neighbors and rainfall on survival.
Survival ($s$) of seedling record *i* of individual *m* for species *j* in census *t* in plot *p* in seedling station *s* was modeled using the Bernoulli distribution ($\mathcal{B}$):

$$
s_{i,j,m,s,t,p} \sim \mathcal{B}(p_{i,j,m,s,t,p}),
$$

$$
\mathrm{logit}(p_{i,j,m,s,t,p}) = \boldsymbol{x_{i}} \cdot \boldsymbol{\beta_{j}} + \eta_m + \phi_s + \psi_p + \xi_t,
$$

where $\boldsymbol{\beta_{j}} = \left[\beta_{1,j}, \beta_{2,j}, \ldots, \beta_{K,j} \right]$ is the coefficient row *K*-vector for species *j*,
*K* is the number of predictors for an individual seedling,
$\boldsymbol{x_i} = \left[x_{i,1},x _{i,2}, \ldots,x_{i,K} \right]$ is the vector of predictors of size *K* for an individual seedling,
$\eta_m$ is the random effect for seedling individual,
$\phi_s$ is the random effect for seedling plots,
$\psi_p$ is the random effect for seedling stations, and
$\xi_t$ is the random effect for different censuses
(note that $\cdot$ denotes dot product).
The set of predictor variables ($\boldsymbol{x_i}$) includes
intercepts,
log of seedling heights,
rainfall,
densities of conspecific (*ConS*) and heterospecific (*HetS*) seedlings,
densities of conspecific (*ConA*) and heterospecific (*HetA*) trees that are scaled by 0.27 for the dry season or 0.24 for the rainy season,
and the interactions of rains with *ConS*, with *HetS*, with *ConA* and with *HetA*.

In the species-level regression, the coefficients ($\boldsymbol{\beta_{1-K}}$) of each species *j* were assumed to have a multivariate normal distribution (MVN),

$$
\boldsymbol{\beta_j} \sim MVN(\boldsymbol{\gamma_k} \cdot \boldsymbol{u_j}, \Sigma),
$$

where $\boldsymbol{u_{j}} = \left[u_{1,j}, u_{2,j}, \ldots, u_{L,j} \right]$ is the row vector of predictors of size *L* for species *j*, with *L* indicaiting the number of predictors for each species, including an intercept,
$\boldsymbol{\gamma_k} = \left[\gamma_{k,1}, \gamma_{k,2}, \ldots, \gamma_{k,L} \right]$ is the coefficient *L*-vector for *k*th predictor in the individual-level regression, and
$\Sigma$ represents the covariance matrix of the coefficients.
We modeled four different sets of species-level predictors separately.
The set of species-level predictor variables ($\boldsymbol{u_j}$) includes
1) leaf and stem traits (
LDMC,
SDMC,
LA,
SLA,
LT,
$\delta$C~13~,
C,
N, and
$\pi$~tlp~),
2) principal components of the traits (PC1 and PC2),
3) abundance (number of individuals), and
4) basal area.
The row vector $\gamma_{k,1}$ represents the average effects of each individual-level predictor (e.g., *ConS*) across species,
whereas $\gamma_{k, l}$ ($l \ne 1$), represents the effects of the *l*-th individual-level predictor (e.g., SLA) on the variation in the strength of each individual-level predictor (e.g., variation in the strength of *ConS* among species).
To allow comparisons among parameter estimates, the individual-level predictors ($\boldsymbol{x_i}$) and the species-level predictors ($\boldsymbol{u_j}$) were scaled to a mean of 0 and a standard deviation of 1 within each season and across species, respectively.
This scaling was applied separately for each season for $\boldsymbol{x_i}$ and across all species for $\boldsymbol{u_j}$.
The complete model details, including the Cholesky factorization and prior information, are provided in Appendix S1.

Posterior distributions of all parameters were estimated using the No-U-Turn Sampler (NUTS), an adaptive variant of the Hamiltonian Monte Carlo (HMC), algorithm implemented in Stan [@Carpenter2017] using the weakly-informative priors [@Gelman2008].
The HMC algorithm uses gradient information to propose new states in the Markov chain, leading to a more efficient exploration of the target distribution than traditional Markov chain Monte Carlo (MCMC) methods that rely on random proposals [@Carpenter2017].
This efficiency allows us to achieve convergence with fewer iterations than traditional MCMC methods.
Four independent chains were run for 2,000 iterations for each model with a warm-up of 1,000 iterations.
Convergence of the posterior distribution was assessed using the Gelman-Rubin statistic with a convergence threshold of 1.1 [@Gelman2013], ensuring effective sample sizes greater than 400 [@Vehtari2021], and by monitoring divergent transitions [@Betancourt2016] for all parameters.

All statistical analyses were conducted in R version 4.3.1 [@RCoreTeam2022] using the R package *targets* version 1.3.2 for workflow management [@Landau2021].
Codes and datasets are available at
<https://github.com/mattocci27/trait-ndd>
and
<https://doi.org/10.57760/sciencedb.02276>,
respectively.

# References
