---
fontsize: 12pt
geometry: margin=1in
link-citations: yes
# csl: templates/ecology-letters.csl
# bibliography: templates/seedling.bib
crossref:
  fig-title: Fig.
  fig-prefix: Fig.
  eq-prefix: Eq.
format:
  html:
    theme: spacelab #readable #sandstone #spacelab #flatly
    toc: true
    toc-depth: 2
    toc-title: Contents
    self-contained: true
    smooth-scroll: true
    highlight-style: github
  docx:
    toc: false
    number-sections: false
    highlight-style: github
    html-math-method: katex
  pdf:
    toc: false
    keep-tex: true
    pdf-engine: pdflatex
    include-in-header:
      text: |
        \usepackage{xr}
        \usepackage[default]{sourcesanspro}
        \usepackage{sourcecodepro}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```



```{r}
library(tidyverse)
library(patchwork)
library(here)
library(kableExtra)
```

```{r}
source(here("R", "stan.R"))
```

```{r, include=FALSE}
dry_gamma <- read_csv(here("data", "dry_gamma.csv"))
dry_gamma |>
  filter(lwr2_5 * upr97_5 > 0)
dry_gamma |>
  filter(trait_name == "intercept")
```

# Table S1

Posterior of distributions of the effects of species-level predcitors on indivdiual-level precitors ($\gamma$) in the dry seasons.
Posterior means, 90% and 95% credible intervals (CI) are shown.
(
ConS/HetS: conspecific/heterospecific seedling density,
ConT/HetT: conspecific/heterospecific tree density,
LDMC: leaf dry matter content,
SDMC: stem dry matter content,
Chl: chlorophyll content,
$\delta$C~13~: stable carbon isotope composition,
C: carbon concentration,
N: nitrogen concentration,
$\pi$~tlp~: leaf turgor loss point,
LA: leaf area,
SLA: specific leaf area,
LT: leaf thickness
)

```{r}
dry_gamma |>
  dplyr::select(-sig) |>
  mutate(pred_name = case_when(
    pred_name == "int" ~ "Intercept",
    pred_name == "logh_scaled" ~ "log[Height]",
    pred_name == "cons_scaled" ~ "ConS",
    pred_name == "cona_scaled_c" ~ "ConT",
    pred_name == "hets_scaled" ~ "HetS",
    pred_name == "heta_scaled_c" ~ "HetT",
    pred_name == "rain_scaled" ~ "Rainfall",
    pred_name == "cons_rain" ~ "ConS $\\times$ Rainfall",
    pred_name == "cona_rain" ~ "ConT $\\times$ Rainfall",
    pred_name == "hets_rain" ~ "HetS $\\times$ Rainfall",
    pred_name == "heta_rain" ~ "HetT $\\times$ Rainfall"
  )) |>
  mutate(trait_name = case_when(
    trait_name == "intercept" ~ "Intercept",
    trait_name == "ldmc" ~ "LDMC",
    trait_name == "sdmc" ~ "SDMC",
    trait_name == "chl" ~ "Chl",
    trait_name == "c13" ~ "$\\delta \\mathrm{C_{13}}$",
    trait_name == "c_mass" ~ "C",
    trait_name == "n_mass" ~ "N",
    trait_name == "tlp" ~ "$\\pi_\\mathrm{{tlp}}$",
    trait_name == "log_la" ~ "log[LA]",
    trait_name == "log_sla" ~ "log[SLA]",
    trait_name == "log_lt" ~ "log[LT]")) |>
  mutate(mean_ = round(mean_, 3)) |>
  mutate(lwr2_5 = round(lwr2_5, 3)) |>
  mutate(lwr5 = round(lwr5, 3)) |>
  mutate(upr95 = round(upr95, 3)) |>
  mutate(upr97_5 = round(upr97_5, 3)) |>
  mutate(`90\\% CI` = paste0("[", lwr5, ", ", upr95, "]")) |>
  mutate(`95\\% CI` = paste0("[", lwr2_5, ", ", upr97_5, "]")) |>
  mutate(para = paste0(
    "$\\gamma_{",
    str_split_fixed(dry_gamma$para, "_" ,3)[,2],
    ",",
    str_split_fixed(dry_gamma$para, "_" ,3)[,3],
    "}$")) |>
  rename(Parameter = para) |>
  rename(Mean = mean_) |>
  rename(`Lower 2.5\\% CI` = lwr2_5) |>
  rename(`Lower 5\\% CI` = lwr5) |>
  rename(`Upper 95\\% CI` = upr95) |>
  rename(`Upper 97.5\\% CI` = upr97_5) |>
  rename(`Individual-level predictor` = pred_name) |>
  rename(`Species-level predictor` = trait_name) |>
  # kable(format = "latex") #|>
  dplyr::select(Parameter, Mean,
  # `Lower 2.5\\% CI`#,
  `90\\% CI`,
  `95\\% CI`,
  `Individual-level predictor`,
  `Species-level predictor`
  ) |>
  kbl(booktabs = TRUE, escape = FALSE, format = "latex", longtable = TRUE) |>
  kable_styling(latex_options = c("striped", "scale_down", "HOLD_position", "repeat_header"))
  #kable_styling()
  #|
  #mutate(hoge = paste0("Effect of ", trait_name, " on ", pred_name))
```
