---
fontsize: 12pt
geometry: margin=1in
link-citations: yes
csl: templates/ecology-letters.csl
bibliography: templates/seedling.bib
crossref:
  fig-title: Fig.
  fig-prefix: Fig.
  eq-prefix: Eq.
format:
  docx:
    toc: false
    number-sections: false
    highlight-style: github
    html-math-method: katex
  html:
    theme: spacelab #readable #sandstone #spacelab #flatly
    toc: true
    toc-depth: 2
    toc-title: Contents
    self-contained: true
    smooth-scroll: true
    highlight-style: github
    embed-resources: true
  pdf:
    toc: false
    keep-tex: true
    pdf-engine: pdflatex
    include-in-header:
      text: |
        \usepackage{xr}
        \usepackage[default]{sourcesanspro}
        \usepackage{sourcecodepro}
        \usepackage{fancyhdr}
        \usepackage{fvextra}
        \pagestyle{fancy}
        \fancypagestyle{plain}{\pagestyle{fancy}}
        \renewcommand{\headrulewidth}{0pt}
        \fancyhead[RE,RO]{Song \textit{et al}. --New Phytologist?-- Appendix X}
        \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


```{r}
library(tidyverse)
library(patchwork)
library(here)
library(kableExtra)
library(flextable)
```

```{r}
source(here("R", "stan.R"))
source(here("R", "render.R"))
```

```{r, include=FALSE}
s_table("dry_traits_diagnostics")
s_table("wet_traits_diagnostics")
s_table("dry_abund_diagnostics")
s_table("wet_abund_diagnostics")
```

```{r, include=FALSE}
dry_traits_gamma <- read_csv(here("data", "dry_het_intrain2_traits_gamma.csv"))
dry_traits_gamma |>
  filter(q2.5 * q97.5 > 0)
dry_traits_gamma |>
  filter(trait_name == "intercept")
```

```{r, include=FALSE}
gen_si_tab2 <- function(data) {
  data2 <- data |>
    mutate(pred_name = case_when(
      pred_name == "(Intercept)" ~ "Intercept",
      pred_name == "logh_s" ~ "ln Height",
      pred_name == "scon_s" ~ "ConS",
      pred_name == "acon_s_c" ~ "ConT",
      pred_name == "shet_s" ~ "HetS",
      pred_name == "ahet_s_c" ~ "HetT",
      pred_name == "sphy_s" ~ "PhyS",
      pred_name == "aphy_s_c" ~ "PhyT",
      pred_name == "rain_s" ~ "Rainfall",
      pred_name == "scon_s:rain_s" ~ "ConS $\\times$ Rainfall",
      pred_name == "acon_s_c:rain_s" ~ "ConT $\\times$ Rainfall",
      pred_name == "shet_s:rain_s" ~ "HetS $\\times$ Rainfall",
      pred_name == "ahet_s_c:rain_s" ~ "HetT $\\times$ Rainfall",
      pred_name == "sphy_s:rain_s" ~ "PhyS $\\times$ Rainfall",
      pred_name == "aphy_s_c:rain_s" ~ "PhyT $\\times$ Rainfall",
      pred_name == "scon_s:rain_s" ~ "ConS $\\times$ Rainfall",
      pred_name == "logh_s:rain_s" ~ "ln Height $\\times$ Rainfall",
    )) |>
    mutate(trait_name = case_when(
      trait_name == "intercept" ~ "Intercept",
      trait_name == "ldmc" ~ "LDMC",
      trait_name == "sdmc" ~ "SDMC",
      trait_name == "chl" ~ "Chl",
      trait_name == "c13" ~ "$\\delta \\mathrm{C_{13}}$",
      trait_name == "log_n" ~ "ln N",
      trait_name == "tlp" ~ "$\\pi_\\mathrm{{tlp}}$",
      trait_name == "log_la" ~ "ln LA",
      trait_name == "log_sla" ~ "ln SLA",
      trait_name == "log_lt" ~ "ln LT",
      trait_name == "log_ab" ~ "ln Abundance",
      )) |>
    mutate_if(is.numeric, round, 3) |>
    filter(q2.5 * q97.5 > 0) #|>
  data2 |>
    mutate(`90\\% CI` = paste0("[", q5, ", ", q95, "]")) |>
    mutate(`95\\% CI` = paste0("[", q2.5, ", ", q97.5, "]")) |>
    mutate(variable = paste0(
      "$\\gamma_{",
      str_split_fixed(data2$variable, ",|\\[|\\]" ,4)[,2],
      ",",
      str_split_fixed(data2$variable, ",|\\[|\\]" ,4)[,3],
      "}$")) |>
    rename(Parameter = variable) |>
    rename(Median = q50) |>
    rename(`Lower 2.5\\% CI` = q2.5) |>
    rename(`Lower 5\\% CI` = q5) |>
    rename(`Upper 95\\% CI` = q95) |>
    rename(`Upper 97.5\\% CI` = q97.5) |>
    rename(`Individual-level predictor` = pred_name) |>
    rename(`Species-level predictor` = trait_name) |>
    dplyr::select(Parameter, Median,
    `95\\% CI`,
    `Individual-level predictor`,
    `Species-level predictor`
    ) |>
    kbl(booktabs = TRUE, escape = FALSE, format = "pipe", longtable = TRUE) |>
    kable_styling(latex_options = c("striped", "scale_down", "HOLD_position", "repeat_header"))
}
```

# `r table("dry_traits_gamma")`

Posterior medians and 95% credible intervals (CIs) of the effects of species-level predictors (traits) on individual-level predictors ($\gamma$) in the dry seasons.
Only effects of those with 95% CIs that did not overlap with zero are shown.
The full results are available in `r s_table("dry_traits_gamma")`.
ConS/HetS: conspecific/heterospecific seedling density,
ConT/HetT: conspecific/heterospecific tree density (scaled by 0.27),
LDMC: leaf dry matter content,
SDMC: stem dry matter content,
$\delta$C~13~: stable carbon isotope composition,
LT: leaf thickness.

```{r}
gen_si_tab2(dry_traits_gamma)
```

# `r table("wet_traits_gamma")`

Posterior medians and 95% credible intervals (CIs) of the effects of species-level predictors (traits) on individual-level predictors ($\gamma$) in the rainy seasons.
Conspecific (ConT) and phylogenetic (PhyT) tree densities are scaled by 0.24.
The full results are available in `r s_table("wet_traits_gamma")`.
$\delta$C~13~: stable carbon isotope composition,
Other details as for `r table("dry_traits_gamma")`.

```{r, include=FALSE}
wet_traits_gamma <- read_csv(here("data", "wet_phy_norain_traits_gamma.csv"))
wet_traits_gamma |>
  filter(q2.5 * q97.5 > 0)
wet_traits_gamma |>
  filter(trait_name == "intercept")
```

```{r}
gen_si_tab2(wet_traits_gamma)
```
